News Topic Modeling

### 2017-11-02 paper/R/03_stm.Rmd
stm

```{r caching, echo = FALSE}
load("../output/models/LDAGibbs 40 .Rda")
load("../output/models/STM 40 .Rda")
#load("../output/btw_combined.Rda")
knitr::opts_chunk$set(eval = TRUE, warning = FALSE)

btw$articleID <- rownames(btw)
```

@@

### LDA Training

```{r eval=FALSE, include=FALSE}
# Build Corpus
btw %>%
  filter(nchar(text_cleaned) > 5) -> btw

# Build corpus
df.lda <- Corpus(VectorSource(btw$text_cleaned))
# Create document-term matrix
dtm <- DocumentTermMatrix(df.lda)
# Assign name to docs
rownames(dtm) <- make.unique(btw$title)

# ---------- STM ----------- #

btw$month <- month(btw$date)
# Process data
processed <- textProcessor(btw$text_cleaned, metadata = btw[,c("date","month","site","text_cleaned","text")],
                           wordLengths = c(2,Inf),
                           lowercase = F,
                           removestopwords = F,
                           removenumbers = F,
                           removepunctuation = F,
                           stem = F)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
```

@@

```{r eval=FALSE, include=FALSE}
## Run LDA using Gibbs sampling

#Set parameters for Gibbs sampling (see https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/)
burnin <- 1000
iter <- 1000
keep <- 50

k = 40 # Topics
```

```{r}
t1 <- Sys.time()
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(burnin=burnin, iter = iter, keep = keep))
t2 <- Sys.time()
t2 - t1

save(ldaOut, btw, dtm, file=paste("../output/models/LDAGibbs",k,".Rda"))

# ---------- STM ----------- #

@@

@@

```{r Extract wtp and dtp}

# Word-topic probabilities
ldaOut %>% tidy("beta") -> posts.wtp

# Document-topic probabilities
ldaOut %>% tidy("gamma") -> posts.dtp

@@

We can then assess the word-topic probabilities in order to get an idea of the topic that is most.

Since we have over 30,000 unique terms in the corpus, we need to extract the top few words that most uniquely define each topic, so that we can more easily visualize and label them. We use the measure of relevance defined by Sievert and Shirley (2014)^[[Sievert and Shirley (2014). LDAvis: A method for visualizing and interpreting topics. *Stanford University.*](http://nlp.stanford.edu/events/illvi2014/papers/sievert-illvi2014.pdf)]. Relevance of term $w$ to topic $k$ given a weight parameter $\lambda$ between 0 and 1 $r(w, k | \lambda)$ is computed as:

$$
r(w, k | \lambda) = \lambda\log(\phi_{kw}) + (1-\lambda)\log(\frac{\phi_{kw}}{p_w})
$$

where $\phi_{kw}$ is the probability of term $w$ for topic $k$, and $p_w$ is the empirical probability of the word in the corpus. $\lambda$ can be thought of as a weighting term between ranking by the probability of that word within the topic, and ranking by the lift over the overall probability in the corpus.

In user studies, Sievert and Shirley (2004) found that a lambda value of 0.6 as an optimal value for allowing humans to identify the topics associated with the top words ranked by relevance. We use this same value for lambda.

```{r Compute relevance}

posts.wtp %>%
  # Compute lambda and phi_kw
  mutate(lambda = 0.6, phi_kw = beta) %>% 
  # Compute and join the p_w
  left_join(
    posts_tokenized.dt %>%
      group_by(word) %>% 
      dplyr::summarise(frequency = sum(term_frequency)) %>% 
      ungroup() %>% 
      mutate(p_w = frequency/sum(frequency)) %>% 
      select(-frequency),
    by = c("term" = "word")
  ) %>%
  # Compute the relevance
  mutate(relevance = lambda * log(phi_kw) + (1 - lambda) * log(phi_kw/p_w)) ->
  word_relevance.dt

```{r Loading topic mappings}
### Most likely Topics per Article

# The topics function from the package is used to extract the most likely topic for each document 
btw.topics <- topics(ldaOut, 2)

# Create Dataframe
doctopics.df <- as.data.frame(t(btw.topics))

doctopics.df %>%
  transmute(title = rownames(.),
            topic = V1) -> doctopics.df

doctopics.df$articleID <- as.integer(rownames(doctopics.df))

# Add Topic to origian DF
topics_mapping.dt <- btw %>%
  mutate(articleID = as.integer(articleID)) %>%
  inner_join(.,doctopics.df, by="articleID") %>%

  group_by(topic) %>%
  slice(1) %>%
  select(topic, title.text, articleID) 
```

```{r}
### Create Topic Labels

## Return the top 30 terms.
btw.terms <- as.data.frame(terms(ldaOut, 30), stringAsFactors = FALSE)

topicTerms <- btw.terms %>% gather(Topic)
topicTerms <- cbind(topicTerms, Rank = rep(1:30))
topicTerms <- topicTerms %>% filter(Rank < 5)

topicTerms <- topicTerms %>% mutate(topic = stringr::word(Topic, 2))
topicTerms$topic <- as.numeric(topicTerms$topic)

@@

set.seed(9272)
posts.dtp %>% 
  group_by(title = document) %>% 
  summarise(topic = min(topic[gamma == max(gamma)]), gamma = max(gamma)) %>% 
  ungroup() %>% 
  inner_join(btw %>% select(title, articleID), by = "title") %>% 
  inner_join(topics_mapping.dt, by = "topic") %>% 
  mutate(topic_title = 
           paste0("Topic ", formatC(topic, flag = "0", width = 2), 
                  " - ", topic_name)) %>% 
  group_by(topic_name) %>% 
  top_n(300, gamma) %>%
  sample_n(5, replace = TRUE) %>% 
  mutate(row = row_number()) %>% 
  ungroup() ->
  posts_classification.sdt

@@

CHART NOTES:
  1. Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics
  2. Documents were classified according to the topic that has the highest probability
"
      )
  },
  device = "png",
  filename = "../figs/02-topic-classification-sample.png",
  dpi = 600, height = 10, width = 16
)

@@

![Facebook news samples](../figs/02-topic-classification-sample.png)

### News landscape

After performing the topic modeling, we can generate an overview of the news landscape about the german elections, each group is composed of the most relevant words for that topic, and a manually-determined label. The distances that the topics have from each other reflect their semantic distance, or basically how different the words are for that topic. It is computed as the Jensen-Shannon divergence between each of the topics according to the differences in their word-topic probabilities. We then apply principal coordinates analysis in order to project the distance matrix down to two dimensions.

```{r echo=TRUE}
posts.wtp %>% 
  # Compute Jensen-Shannon
  computeJensenShannonPCA() %>% {

    # Assign positions to top 30 most relevant words
    left_join(
      .,
      word_relevance.dt %>% 
        group_by(topic) %>% 
        top_n(30, relevance) %>% 
        ungroup(),
      by = "topic"
    ) -> topic_labelling_top30words.dt

    # Assign positions to topic labels
    left_join(
      .,
      topics_mapping.dt,
      by = "topic"
    ) -> topic_labelling_topics.dt

    ggsave(
      plot = {
        topic_labelling_top30words.dt %>% 
          ggplot(aes(x = x, y = y)) +
          geom_text_repel(
            aes(label = term, 
                color = factor(topic, levels = 1:40), 
                size = phi_kw * p_w), 
            lineheight = 0.6, 
            segment.size = NA,
            box.padding = unit(0, "lines"),
            force = 0.05,
            family = "Roboto"
          ) +
          geom_text_repel(
            data = topic_labelling_topics.dt,
            aes(label = str_wrap(topic_name, 15),
                color = factor(topic, levels = 1:40)),
            size = 2.25, lineheight = 0.7,
            segment.size = 0.5, alpha = 0.7,
            family = "Roboto", fontface = 2,
            force = 0.3
          ) +
          scale_size_continuous(range = c(1, 4)) +
          scale_color_manual(
            values = rainbow(40) %>% 
              adjustcolor(red.f = 0.6, green.f = 0.6, blue.f = 0.6)
          ) +
          theme(
            axis.title = element_blank(),
            axis.text  = element_blank(),
            axis.ticks = element_blank(),
            panel.grid = element_blank(),
            panel.border = element_blank(),
            plot.background = element_rect(fill = "#FAFAFA"),
            legend.position = "none"
          ) +
          labs(
            title    = "GERMAN NEWS LANDSCAPE",
            subtitle = "Online News Articles about Federal Elections June-Oct 2017",
            caption = 
"DATA SOURCE: Webhose.io, Eventbride.com

CHART NOTES:
  1. Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics
  2. For each topic, the top 30 words in terms of relevance as defined in Sievert and Shirley (2014) are displayed
  3. The centers of each topic mass were determined by applying principal components analysis to the distance matrix, with the Jensen-Shannon divergence as the distance measure. 
      This layout is intended to display the degree of semantic distance.
"
          )
      },
      filename = "../figs/01-news-landscape-map.png",
      device = "png",
      dpi = 600
    )

  }

![Facebook News Map](../figs/01-news-landscape-map.png)

### Topic Trends

We classify each document based on which topic has the highest probability. If one or more topics have equal probability, then the count is assigned equally to all the topics.

```{r Topic classification}

posts.dtp %>% 
  group_by(document) %>% 
  filter(gamma == max(gamma)) %>% 
  mutate(allocation = 1/n()) %>% 
  ungroup() ->
  posts_classification.dt

We construct a chart that allows us to explore topic trends, both in terms of document count and the total number of reactions on the articles of that particular topic.

```{r Data for topic trends chart}

set.seed(7292)
posts_classification.dt %>% 
  # Add post information
  inner_join(
    btw %>% 
      select(document = title, date, 
             title, articleID) %>% 
      mutate(post_period = 
               date %>% 
               with_tz("Europe/Paris") %>% 
               floor_date("week")),
    by = "document"
  ) %>%
  # Add topic information
  left_join(topics_mapping.dt, by = "topic") %>%
  # Summarise into tidy dataframe
  group_by(post_period, topic_name) %>% 
  summarise(
    articles = sum(allocation),
    sample_articles = 
      paste0("<li>", sample(document, 3, replace = TRUE ), "</li>", collapse = "<br>")
  ) %>% 
  ungroup() %>% 
  mutate(
    tooltip_text = 
      paste0(
        "<b>", topic_name, "</b> in period of <b>", post_period, "</b><br>",
        "<b>Number of Articles:</b> ", comma(round(articles, 0)), "<br>",
        "<b>Sample Articles:</b><br><ul>", sample_articles, "</ul>"
      )
  ) %>% 
  # Mark peaks 
  group_by(topic_name) %>% 
  mutate_at(
    vars(articles), 
    funs(mean = mean(.), sd = sd(.), zscore = (. - mean(.))/sd(.))
  ) %>% 
  ungroup() %>% 
  mutate_at(
    vars(ends_with("zscore")),
    funs(peak = . >	1.96)
  ) %>% 
  mutate(tooltip_text = str_replace_all(tooltip_text, "'", "")) ->
  chart_topic_trends.dt
```

```{r Topic trends chart, eval = TRUE}

{
  chart_topic_trends.dt %>% 
    ggplot(
      aes(
        x = post_period, 
        y = factor(
          topic_name, 
          chart_topic_trends.dt %>% 
            group_by(topic_name) %>% 
            summarise(post_period = mean(post_period[peak], na.rm = TRUE)) %>% 
            ungroup() %>% 
            arrange(desc(post_period)) %$%
            unique(topic_name)
        )
      )
    ) +
    geom_tile_interactive(aes(tooltip = tooltip_text, fill = log(articles))) +
    geom_point_interactive(
      mapping = aes(tooltip = tooltip_text),
      data = chart_topic_trends.dt %>% filter(peak),
      color = "red"
    ) +
    scale_fill_viridis(
      labels = function(x) comma(round(exp(x))), 
      name = "Number of Articles",
      option = "plasma"
    ) +
    scale_x_datetime(expand = c(0, 0)) +
    labs(
      title = "ZEITGEIST", 
      subtitle = "Topic distribution of news articles over time, 2017\n(Hover over each tile to see sample articles)",
      caption = "
DATA SOURCE: Webhose.io, Eventbride.com

CHART NOTES:
  1. Red dots indicate weeks where the number of articles exceeds 1.96 SD above mean for that topic
  2. Hover over each tile to see infomration and to see sample articles
  3. Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics
  4. Articles were classified according to the topic that has the highest probability

TROY JAMES PALANCA | TJPALANCA.COM
      "
    ) +
    theme(
      legend.position   = "bottom",
      legend.key.width  = unit(0.1, "npc"),
      legend.background = element_rect("#fafafa"),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      panel.grid        = element_blank(),
      panel.background  = element_rect("#fafafa"),
      axis.title        = element_blank(),
      axis.ticks        = element_blank(),
      axis.text.y       = element_text(size = 8)
    ) 
} %>% 
  ggiraph(
    ggobj = .,
    # width_svg  = 8,
    # height_svg = 7,
    width = 1,
    tooltip_extra_css = "
    font-family: Roboto; 
    background-color: #000; 
    color: #fff; font-size: 10px;
    padding: 5px;"
  ) %>% 
  prependContent(
    tags$link(
      href = "https://fonts.googleapis.com/css?family=Roboto:400,700", 
      rel = "stylesheet"
    )
  ) -> chart_topic_trends.wdgt

saveWidget(
  widget = chart_topic_trends.wdgt,
  file = "../figs/03-topic-trends.html",
  selfcontained = FALSE,
  libdir = "../figs/js",
  background = "#fafafa"
)

chart_topic_trends.wdgt

```{r Topic trends line chart, eval = TRUE}

{ 
  chart_topic_trends.dt %>%
  select(post_period, topic_name, articles) %>% 
  mutate(data_id = row_number() %>% as.character()) %>% 
  ggplot(aes(x = post_period, y = articles)) +
  geom_path_interactive(
    aes(group = topic_name, 
        tooltip = str_replace_all(topic_name, "'", ""),
        data_id = data_id), 
    color = "gray"
  ) +
    scale_y_continuous(labels = comma) +
    labs(
      title = "PEAKS AND TROUGHS",
      subtitle = "Hover over each line to see the trend of news articles for a particular topic",
      y = "Number of Articles"
    ) +
    theme(
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      panel.grid        = element_blank(),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )
} %>% 
  ggiraph(
    ggobj = .,
    hover_css = "stroke: red; stroke-width: 3px;",
    width = 1,
    width_svg = 8,
    height_svg = 4,
    tooltip_extra_css = "
    font-family: Roboto; 
    background-color: #000; 
    color: #fff; font-size: 10px;
    padding: 5px;"
  ) %>% 
  prependContent(
    tags$link(
      href = "https://fonts.googleapis.com/css?family=Roboto:400,700", 
      rel = "stylesheet"
    )
  ) -> chart_trends_line.wdgt

saveWidget(
  widget = chart_trends_line.wdgt,
  file = "../figs/06-topic-trends-line.html",
  selfcontained = FALSE,
  libdir = "../figs/js",
  background = "#fafafa"
)

chart_trends_line.wdgt

```{r Trend to reactions, eval = TRUE}

# Prepare dataset for joy plot
posts_classification.dt %>% 
  inner_join(
    btw %>% 
      select(
        document = title, 
        news_page = site, 
        title, 
        date
      ),
    by = "document"
  ) %>% 
  inner_join(topics_mapping.dt, by = "topic") %>% 
  inner_join(
    btw %>% 
      group_by(document = title) %>%
      summarise(reactions = sum(fb_shares)) %>%
      ungroup(),
    by = "document"
  ) %>%
  group_by(
    topic_name, 
    # 8 hour time difference (Manila)
    week = floor_date(date, "week")
  ) %>% 
  summarise(
    num_articles  = sum(allocation),
    num_reactions = sum(reactions * allocation)
  ) %>%
  #mutate(peak = week[num_articles == max(num_articles)]) %>%
  ungroup() %>%
  arrange(week) %>%
  mutate(
    topic_name = factor(topic_name, levels = rev(unique(topic_name))),
    # ID needed for linking back information after tweening frames
    id = as.character(row_number()),
    # Maximum height for ridgeline is 9
    articles_height = num_articles/max(num_articles) * 9,
    reactions_height = num_reactions/max(num_reactions) * 9
  ) ->
  posts_time_trend.dt

# Prepare animated dataset
bind_rows(
  # Articles
  posts_time_trend.dt %>% 
    transmute(id, time = 0,  ht = articles_height , op = 0),
  # Reactions
  posts_time_trend.dt %>% 
    transmute(id, time = 10, ht = reactions_height, op = 1),
  # Return to articles
  posts_time_trend.dt %>% 
    transmute(id, time = 20, ht = articles_height , op = 0)
) %>%
  # Add cubic easing
  mutate(ease = "cubic-in-out") %>%
  tween_elements("time", "id", "ease") %>%
  # Add back the topic name and week
  inner_join(
    posts_time_trend.dt %>% select(topic_name, week, id), 
    by = c(".group" = "id")
  ) ->
  posts_time_trend.adt

# Create labels above the chart (what news outlets post vs what we react to)
posts_time_trend.adt %>%
  group_by(.frame) %>%
  summarise(
    opacity    = mean(op),
    range_week = max(week) - min(week),
    min_week   = min(week),
    num_topics = n_distinct(topic_name)
  ) %>% 
  ungroup() -> 
  posts_time_trend_labels.adt

```{r Trend to reactions plot, eval = TRUE}

# Create animated pot
posts_time_trend.adt %>%
  ggplot(aes(x = week, y = topic_name, height = ht, 
             fill = topic_name, frame = .frame)) +
  geom_ridgeline(color = NA, alpha = 0.5) +
  geom_text(
    data = posts_time_trend_labels.adt,
    aes(x = min_week + 0.25 * range_week, 
        y = num_topics + 3.5, 
        alpha = 1 - opacity, 
        frame = .frame),
    label = "What news outlets publish", color = "forestgreen",
    family = "Roboto", fontface = "bold", size = 9,
    inherit.aes = FALSE
  ) +
  geom_text(
    data = posts_time_trend_labels.adt,
    aes(x = min_week + 0.75 * range_week, 
        y = num_topics + 3.5, 
        alpha = opacity, 
        frame = .frame),
    label = "What we react to", color = "firebrick",
    family = "Roboto", fontface = "bold", size = 9,
    inherit.aes = FALSE
  ) +
  geom_text(
    data = posts_time_trend_labels.adt,
    aes(x = min_week + 0.55 * range_week, 
        y = num_topics + 3.5, 
        frame = .frame),
    label = "VS",
    family = "Roboto", fontface = "bold", size = 7,
    inherit.aes = FALSE
  ) +
  scale_x_datetime(expand = c(0, 0)) +
  theme(
    legend.position = "none",
    panel.border    = element_blank(),
    axis.title      = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.caption      = element_text(size = 9)
  ) + 
  labs(
    caption = "
DATA SOURCE: Webhose.io, Eventbride.com
CHART NOTE: Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics for online news articles in 2017 containing the term 'Bundestagswahl'
    "
  ) ->
  posts_time_trend.agg

gganimate(
  p = posts_time_trend.agg,
  filename = "../figs/09-reactions-vs-articles.gif",
  title_frame = FALSE,
  interval = 1/10,
  ani.width = 900,
  ani.height = 600
)

![Articles to Reactions](../figs/09-reactions-vs-articles.gif)

### Topic Distribution by News Page

#### Distribution

A common accusation leveled against traditional news media is the amount of bias in reporting various topics. We try to explore the topic distribution of news articles.

```{r Data for topic distribution chart}

set.seed(7292)
posts_classification.dt %>% 
  inner_join(
    btw %>% select(document = title, news_page = site, title),
    by = "document"
  ) %>%
  inner_join(topics_mapping.dt, by = "topic") %>% 
  group_by(news_page, topic_name) %>%
  summarise(
    articles = sum(allocation),
    sample_articles = 
      paste0(
        "<li>", sample(document, pmin(3, length(document))), "</li>",
        collapse = "<br>"
      )
  ) %>%
  ungroup() %>% 
  group_by(news_page) %>% 
  mutate(perc_articles = articles/sum(articles)) %>% 
  ungroup() %>%
  # left_join(
  #   top_100_fbpages.dt %>% select(news_page = page_name, page_title), 
  #   by = "news_page"
  # ) %>% 
  mutate(
    tooltip_text = 
      paste0(
        "<b>", comma(round(articles, 0)), " (", percent(perc_articles), 
        ") </b> of articles from <b>", news_page, 
        "</b> fall under <b>", topic_name, "</b><br><br>Sample Articles:<br><ul>",
        sample_articles, "</ul>"
      ) %>% str_replace_all("'", "")
  ) ->
  news_page_distribution.dt
```

```{r Topic distribution chart, eval = TRUE}
{
  news_page_distribution.dt %>% 
    ggplot(aes(x = news_page, y = perc_articles, fill = topic_name)) +
    geom_bar_interactive(
      aes(tooltip = tooltip_text),
      position = position_stack(), 
      stat = "identity", color = "darkgray"
    ) +
    scale_y_continuous(
      name = "Percent of Articles", 
      labels = percent,
      expand = c(0, 0)
    ) +
    theme(
      legend.position = "none",
      axis.title.y = element_blank(),
      legend.background = element_rect("#fafafa"),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      plot.title        = element_text(size = 18)
    ) +
    coord_flip() +
    scale_fill_viridis(discrete = TRUE) +
    labs(
      title = "BIAS AND BALANCE",
      subtitle = "Topic distribution of online news articles per news page, 2017\n(Hover over each bar to see the topic and sample articles)",
      caption = "
DATA SOURCE: Facebook Graph API

CHART NOTES:
  1. The news pages are arranged in descending order in terms of number of articles
  2. Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics
  3. Articles were classified according to the topic that has the highest probability
      "
    )
} %>% 
  ggiraph(
    ggobj = .,
    width_svg = 12,
    height_svg = 8,
    width = 1,
    tooltip_extra_css = "
    font-family: Roboto; 
    background-color: #000; 
    color: #fff; font-size: 10px;
    padding: 5px;"
  ) %>% 
  prependContent(
    tags$link(
      href = "https://fonts.googleapis.com/css?family=Roboto:400,700", 
      rel = "stylesheet"
    )
  ) -> chart_topic_distribution.wdgt

saveWidget(
  widget = chart_topic_distribution.wdgt,
  file = "../figs/04-topic-distribution-pages.html",
  selfcontained = FALSE,
  libdir = "../figs/js",
  background = "#fafafa"
)

chart_topic_distribution.wdgt

#### Topic Concentration

A common criticism leveled against news media sites is that there is undue emphasis on a particular topic. One way to measure concentration across different topics is the Herfindahl-Hirschman Index (HHI) which is simply $\Sigma_n{s^2}$ where $s$ is the share of each topic.

```{r Topic concentration chart, eval = TRUE}

{
  news_page_distribution.dt %>% 
    group_by(news_page) %>% 
    arrange(desc(articles)) %>% 
    summarise(
      hh_index  = sum(perc_articles ^ 2),
      t3_topics = paste0(topic_name %>% head(3), collapse = "<br>")
    ) %>% 
    ungroup() %>% 
    arrange(hh_index) %>% 
    mutate(news_page = factor(news_page, unique(news_page))) %>% 
    ggplot(aes(x = news_page, y = hh_index * 100)) +
    geom_bar_interactive(
      aes(
        tooltip = 
          paste0("<b>Top Topics:</b><br>", t3_topics) %>% 
          str_replace_all("'", "")
      ),
      stat = "identity", 
      fill = c("#008B45")
    ) +
    geom_text(aes(label = comma(round(hh_index * 100, 1))), hjust = 1.5, 
              color = "white", family = "Roboto") +
    scale_y_log10(
      name = "Herfindahl-Hirschman Index (HHI)\n(100 = most concentrated)",
      expand = c(0, 0)
    ) +
    labs(
      title = "TOPIC CONCENTRATION OF PHILIPPINE NEWS PAGES",
      subtitle = "Herfindahl-Hirschman Index (HHI)",
      caption = "
DATA SOURCE: Facebook Graph API

CHART NOTES:
  1. Hover over each tile to see infomration and to see top topics
  2. Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics
  3. Articles were classified according to the topic that has the highest probability

TROY JAMES PALANCA | TJPALANCA.COM
      "
    ) +
    coord_flip() +
    theme(
      axis.title.y       = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.subtitle      = element_text(size = 10),
      plot.caption       = element_text(size = 8),
      plot.title         = element_text(size = 16)
    )
} %>% 
  ggiraph(
    ggobj = .,
    width_svg = 10,
    height_svg = 8,
    width = 1,
    tooltip_extra_css = "
    font-family: Roboto; 
    background-color: #000; 
    color: #fff; font-size: 16px;
    padding: 5px;"
  ) %>% 
  prependContent(
    tags$link(
      href = "https://fonts.googleapis.com/css?family=Roboto:400,700", 
      rel = "stylesheet"
    )
  ) -> chart_topic_concentration.wdgt

saveWidget(
  widget = chart_topic_concentration.wdgt,
  file = "../figs/05-topic-concentration.html",
  selfcontained = FALSE,
  libdir = "../figs/js",
  background = "#fafafa"
)

chart_topic_concentration.wdgt

>>>>

```{r caching, echo = FALSE}
#load("../output/models/LDAGibbs 40 .Rda")
load("../output/models/STM 40 .Rda")
#load("../output/btw_combined.Rda")
knitr::opts_chunk$set(eval = TRUE, warning = FALSE)

btw$articleID <- as.integer(rownames(btw))
```

@@

### STM Training

```{r eval=FALSE, include=FALSE}
# Build Corpus
btw$month <- month(btw$date)
# Process data
processed <- textProcessor(btw$text_cleaned, metadata = btw[,c("date","month","site","text_cleaned","text")],
                           wordLengths = c(2,Inf),
                           lowercase = F,
                           removestopwords = F,
                           removenumbers = F,
                           removepunctuation = F,
                           stem = F)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
```

@@

```{r}
k = 40 # Topics

@@

### STM Exploratiom

#### Topic Summary
```{r}
plot(stmOut, type="summary", xlim = c(0,.14), n=5, text.cex = 0.7, main = "Topics in german online news covering the Federal Elections")
```

Assign labels to the topics based in the interpretation of word-topic probabilities.

```{r}
Labelprep <- labelTopics(stmOut)

# Topic Labels
topicLabels <- as.data.frame(Labelprep$topics)
topicLabels %>%
  transmute(labels = paste(V1,V2,V3, sep = ", "),
            topic = rownames(.)) -> topicLabels

# Covariate topic labels
rown <- as.data.frame(rep(rownames(Labelprep$topics),6))
rown <- rown %>% 
  transmute(topics = as.numeric(.[,1])) %>% arrange(topics) %>%
  mutate(site = rep(rownames(Labelprepr$covariate), 40))

covLables <- as.data.frame(Labelprepr$interaction)
covLables <- covLables %>%
  transmute(topic = as.numeric(rown$topics),
            site = rown$site,
    labels = paste(V1,V2,V3, sep = ", "),
            topic.site = paste0("Topic", rown$topics, ", ",rown$site))
```

We can also plot the exclusivity and semantic coherence, which are two metrics that measure the “interpretability” of each topic. Higher semantic coherence indicate topics that have more consistent words (more interpretable) while exclusivity measures how exclusive the words are to the topic relative to other topics (e.g. low values mean topics that are vague and share a lot of words with other topics while high values indicate words that are very unique/exclusive to the topic).

Does not support models with content covariates.

```{r}
# par(mfrow = c(1,1),mar = c(2, 2, 2, 2))
# 
# topicQuality(stmOut,
#              documents = out$documents, 
#              main = "Topic Interpretability: Exclusivity and Semantic Coherence")
```

#### Estimating the Effect of covariates
One of the main contributions of STM is to provide a mechanism to test (with statistical significance) the impact document-level covariates (e.g. party, time of post) have on the topic proportions (prevalence). Method used for plotting. "pointestimate" estimates mean topic proportions for each value of the covariate.

```{r Estimating the Effect of Covariates}
k <- 40

prep <- estimateEffect(1:k ~ site+s(month), stmobj = stmOut, meta = out$meta, uncertainty = "Global")
```

```{r Plot Effect Of Covariates On Topics}
par(mfrow = c(3,3), mar = c(4,4,2,2))
for (i in 1:k){
  plot(prep, "site", method = "pointestimate", topics = i, 
       labeltype = "custom", custom.labels = prep$data$site,
       cex = 0.6, cex.main = 0.7,
       main = topicLabels[which(topicLabels$topic==i),])
}
```

We can also explore the effect of a contiuous covariate like time. In this case, we use month using a b-spline to smooth out our results.

For simplicity, I removed confidence intervals and only plotted the (point) estimates.

```{r}
col = palette(rainbow(6))

par(mfrow = c(2,2),mar = c(4,4,2,2))
for (i in 1:k){
  plot(prep, "month", method = "continuous", topics = i, 
                      main = paste0(topicNames$labels[i],"\n: Topic ", topicNames$topic[i]),
                      cex.main = .8,
                      printlegend = FALSE, ylab = "Exp. Topic Prob", 
                      xlab = "Time (June 2017 to Oct 2016)",
                      ylim = c(-0.07,0.35),
                      cex.axis = .6,
                      cex.label = .6,
                      moderator = "site", moderator.value = prep$data$site[1],  linecol = col[1],
                      ci.level = 0
                      )
  for (j in 2:6){
    plot(prep, "month", method = "continuous", topics = i, 
                      #main = paste0(topicNames$labels[i],": Topic ", topicNames$topic[i]),
                      printlegend = FALSE, ylab = "Exp. Topic Prob", 
                      ylim = c(-0.07,0.35),
                      moderator = "site", moderator.value = prep$data$site[j],  linecol = col[j],
                      ci.level = 0, add="T"
                      )
  }

legend(-1, 1.9, prep$data$site, lwd = 2, col = col)
}

prep$parameters
```

@@

```{r Extract wtp and dtp}
# Word-topic probabilities
stmOut %>% tidy("beta") %>% filter(!is.na(topic)) -> posts.wtp

# Document-topic probablities
stmOut %>% tidy("gamma") -> posts.dtp
```

We can then assess the word-topic probabilities in order to get an idea of the topic.

Since we have over 30,000 unique terms in the corpus, we need to extract the top few words that most uniquely define each topic, so that we can more easily visualize and label them.

```{r}
### Create Topic Labels
btw.terms <- labelTopics(stmOut,1, n=30)
btw.terms <- as.data.frame(t(btw.terms$topics))

topicTerms <- btw.terms %>% gather(topic)
topicTerms <- cbind(topicTerms, Rank = rep(1:30))
topicTerms <- topicTerms %>% 
  filter(Rank < 5) %>%
  mutate(topic = as.numeric(topic))

```{r Compute relevance}
word_relevance.dt <- posts.wtp %>%
  filter(str_length(term) >= 3) %>%
  # Compute lambda and phi_kw
  mutate(lambda = 0.6, phi_kw = beta) %>% 
  # Compute and join the p_w
  left_join(
    posts_tokenized.dt %>%
      group_by(word) %>% 
      summarise(frequency = sum(term_frequency)) %>% 
      ungroup() %>% 
      mutate(p_w = frequency/sum(frequency)) %>% 
      select(-frequency),
    by = c("term" = "word")
  ) %>%
  # Compute the relevance
  mutate(relevance = lambda * log(phi_kw) + (1 - lambda) * log(phi_kw/p_w))
  filter(!is.na(relevance))

@@

```{r Loading topic mappings}
### Most likely Topics per Article
btw.topics <- make.dt(stmOut)
btw.topics %>% select(-docnum) -> btw.topics

# Create Dataframe
btw.topics$V1 <- colnames(btw.topics)[apply(btw.topics,1,which.max)]

btw.topics %>%
  select(V1) %>% 
  transmute(articleID = as.integer(rownames(.)),
            topic = as.integer(gsub("Topic","",V1))) -> doctopics.df

# Add Topic to origian DF
btw.dt <- btw %>%
  inner_join(.,doctopics.df, by="articleID")

topics_mapping.dt<-btw.dt %>%
  group_by(topic) %>%
  slice(1) %>%
  select(topic, title.text, articleID, fb_shares) 
```

@@

set.seed(9272)

posts_classification.sdt <- posts.dtp %>% 
  group_by(articleID = document) %>% 
  summarise(topic = min(topic[gamma == max(gamma)]), gamma = max(gamma)) %>% 
  ungroup() %>%
  inner_join(btw %>% select(title, title.text, articleID), by = "articleID") %>%
  inner_join(topics_mapping.dt %>% select(topic, topic_name), by = "topic") %>%
  mutate(topic_title = 
           paste0("Topic ", formatC(topic, flag = "0", width = 2), 
                  " - ", topic_name)) %>%
  group_by(topic_name) %>%
  top_n(300, gamma) %>%
  sample_n(5) %>% 
  mutate(row = row_number()) %>% 
  ungroup() 
```

@@

CHART NOTES:
  1. Topics were derived through Latent Dirichlet Allocation (LDA) with 40 topics
  2. Documents were classified according to the topic that has the highest probability
"
      )
  },
  device = "png",
  filename = "../figs/02-stm-topic-classification-sample.png",
  dpi = 600, height = 10, width = 16
)

@@

![Facebook news samples](../figs/02-stm-topic-classification-sample.png)
